---
title: "HOME WORK 1"
author: "RUTH CHIKA UZOKA"
date: "2023-08-26"
output: pdf_document

****QUESTION 1

WHY DOES THIS CODE NOT WORK?

my_variable <- 10
my_varlable

Look carefully!(This may seem like an exercise in pointlessness, but training your brain to notice even the tiniest difference will pay off when programming.)
```{r }
#ANSWER 1-

my_variable <- 10
my_variable

#initially,The code did not work;-REASON- because of the letter "L" inbetween R and A in the world VARIABLE on the second line. In other words, Variable was mis-spelt as (varlable)and i corrected it on the answer space as VARIABLE before running it and it ran correctly.

```
****QUESTION 2

Tweak each of the following R commands so that they run correctly:
  
library(tidyverse)


ggplot(dota =mpg)+
  geom_point(mapping = aes(x = displ, y = hwy))
  
fliter(mpg, cyl = 8)
filter(diamond, carat > 3)


```{r}
#ANSWER 2
library(tidyverse)

ggplot(data =mpg)+
'geom_point(mpg = aes(x= disp, y= hwy))"

filter(mpg, cyl ==8)
filter(diamonds,carat > 3)

#correction 1-I changed the ggplot function -Dota(incorrect spelling)to data.

#correction 2- I had to use the single quotation mark at the start of the geom_point(enablings layers of plot)and a double quotation mark at the end to run the geom command on R.

#correction 3-I figured there is no such function as Fliter on R but there is an existing function as Filter. Hence the change of the word from FLITER TO Filter Also,the equalisation sign on R, as a comparison operator needs to be doubled to give a functional value. so, i had to double it to run.

#correction 4-Since carat is a word for unit measurement of diamonds, I had to add (S) which was initially missing to correct the word diamonds before running it.
```
*****QUESTION 3

Press Alt + shift+ k. what happens? How can you get to the same place using the menus?

```{r}
#ANSWER 3
After pressing alt+shift+k keys simultaneously,it poped up the system short cut keys..The same function can be called up on R when we go on the "TOOLS"icon in the the top left corner on R studio. it is actually the 10th feature on the R menu,click and go to the "keyboard shortcuts help" option from the drop down.It  gives the same result but different methods to navigate, especially on R. 


```