---
title: "9/27/2023"
author: "UZOKA"
date: "2023-09-20"
output: pdf_document
---
***Evidence of Data Fabrication

1. Let’s compare the distribution of feel in the CCAP dataset to the distribution of feel1 in the
gay marriage experiment and figure out whether Broockman et al. were right when describing
the two distributions as indistinguishable from each other. As you may recall, when comparing
distributions with different number of observations, it is best to create density histograms, so
go ahead and create (1) the density histogram of feel in the CCAP dataset and (2) the density
histogram of feel1 in the gay marriage experiment dataset. (Optional: If you want to ensure that
both histograms have the same number of bins (let’s say 10) so that it is easier to compare them
to each other, we can set the optional argument breaks to equal 10 inside the function hist().)
Do these two distributions look similar to each other? A yes/no answer will suffice.
```{r }
gay <- read.csv("gay.csv")
CCAP <- read.csv("CCAP.csv")
head(CCAP)
head(gay)
library(tidyverse)



CCAP_dens <- ggplot(data = CCAP)
CCAP_dens <- CCAP_dens + geom_histogram(mapping = aes(y = ..density.., x= feel), binwidth=10)
CCAP_dens <- CCAP_dens + theme(aspect.ratio =1) + ylab("Density")  + xlab("Reaction to gay people, CCAP")
CCAP_dens




gay_dens <- ggplot(data = gay)
gay_dens <- gay_dens + geom_histogram(mapping = aes(y =..density.., x = feel1), binwidth=10)
gay_dens <- gay_dens + theme(aspect.ratio =1)  + ylab("Density")  + xlab("Reaction to gay people, gay")
gay_dens



```

From the results on the histogram density plot; Yes, The two distribution looks similar.

2. To further compare the two distributions above, compute the mean, median, and standard deviation
of both. Do the two distributions have similar descriptive statistics? 
```{r}
gay <- read.csv("gay.csv")
CCAP <- read.csv("CCAP.csv")

mean(CCAP$feel) #58.09241
mean(gay$feel1)#58.62884

median(CCAP$feel)#52
median(gay$feel1)#52


sd(CCAP$feel)#28.40383
sd(gay$feel1)#28.55804



```
From the statistical result, yes, The descriptive statistics are similar.

3. Given that the CCAP data is based on a national sample of U.S. voters and that the gay marriage
experiment data was supposedly collected from a sample of residents from the Los Angeles county,
would you have expected that the two distributions of feelings towards gay men and lesbian
women were similar to each other?

*Answer*- No. On average,I would expect that the result should differ because LOS angeles is just a subset(one of states in the U.S)Therefore, the feelings should not be similar for both computations. reason is, why one focuses on the general populace the other focuses on peoples feelings or reactions in just one component(state) of the whole(national populace.

4. As evidence that the data from the gay marriage experiment was not collected as claimed
by LaCour but rather fabricated, Broockman et al. pointed out that it could have been easily
generated by a computer by:

(a) drawing a random sample of observations from the CCAP dataset to produce the baseline
feelings (feel1), which would explain the similarities between the two distributions observed
in questions 1 and 2 above, and then

(b) adding random noise to feel1 to produce the follow-up waves (feel2, feel3, and feel4), which
would explain the unusual high levels of stability in respondent’s feelings towards gay men
and lesbian women over time found in the gay marriage experiment.
####

To explore the stability of respondent’s feelings in the gay marriage experiment data
over time, create the following three visualizations:

(a) the scatter plot of feel1 and feel2 (to show the relationship between respondents feelings
before canvassers went into the field and three weeks after),

(b) the scatter plot of feel1 and feel3 (to show the relationship between respondents feelings
before canvassers went into the field and six weeks after),

(c) the scatter plot of feel1 and feel4 (to show the relationship between respondents feelings
before canvassers went into the field and 9 months after).

```{r }
gay <- read.csv("gay.csv")
CCAP <- read.csv("CCAP.csv")
head(CCAP)
head(gay)

No_gay_tmt <- gay[gay$treatment=="No Contact",] 
 
plot(No_gay_tmt$feel1, No_gay_tmt$feel2)

plot(No_gay_tmt$feel1, No_gay_tmt$feel3)

plot(No_gay_tmt$feel1, No_gay_tmt$feel4)


```
##From the observation in the scattered plot, Yes. There is no significant difference in the response feelings despite that it was conducted at some intervals. This looks suspicious and fabricated because ideally, given that the observation was conducted at 3 weeks interval for each, one would expect that the responses would differ and not look almost stable or somewhat similar.

5. To further explore the stability of respondent’s feelings over time, calculate the following correlations coefficients:
(a) the correlation between feel1 and feel2 (to measure the direction and strength of the linear
association between respondents feelings before canvassers went into the field and three
weeks after),
(b) the correlation between feel1 and feel3 (to measure the direction and strength of the linear
association between respondents feelings before canvassers went into the field and six
weeks after),
(c) the correlation between feel1 and feel4 (to measure the direction and strength of the linear
association between respondents feelings before canvassers went into the field and 9 months
after)
```{r}
gay <- read.csv("gay.csv")

cor(gay$ feel1, gay$feel2, use = "complete.obs" )#0.9608118
cor(gay$ feel1, gay$feel3, use = "complete.obs" )#0.9495854
cor(gay$ feel1, gay$feel4, use = "complete.obs" )#0.9538741


```
##From the results, The variables are correlated because approximately,the answers are close to 1.

6. Given your answers to questions 1 through 5, would you conclude that it is likely that the gay
marriage experiment data was fabricated and not collected as described in the article? A yes/no
answer will suffice.

Observationally, the computational outcome from questions one through 5, The data sets looks fabricated and this is evident in that there is no significant variation in the results. Highly suspicious is that Michael etall must have computerised  random numbers with a specific baseline to generate the variables with slight difference. also, given that Human opinions are subject to change, It is odd to have similar responses after a three weeks interval thrice.


**Predicting Course Grades Using Midterm Scores-Part I: Predicting Final Exam Scores**

1. First, let’s figure out what each observation represents, identify our X and Y variables, and
explore whether they are moderately or strongly linearly associated with each other.

a. In this dataset, what does each observation represent? (2.5 points)

b. What should be our X variable? In other words, which variable are we going to use as the
predictor? Please provide the name of the variable and identify whether it is binary or
non-binary. 

c. What should be our Y variable? In other words, which variable are we going to use as the
outcome variable? Please provide the name of the variable and identify whether it is binary or non-binary. 

d. Compute the correlation coefficient between X and Y. Is the relationship between X and Y
moderately or strongly linear? A yes/no answer will suffice. 

###Answer-

a)From the "grades" dataset,Each observation represents varying students performance in a class.

b)The predictor variable also known as Independent variable is the *Midterm* variable. it is non binary as the numbers ranges from 0 -100

c)The y variable or dependent variable is the *Final* variable. It is non- binary.

d)The correlation coefficient between the midterm variable and the Final variable is 0.7  which indicate a strong positive linear relationship.

```{r}
grades <- read.csv("grades.csv")
head(grades)

corr_l_xy <- cor(grades$midterm, grades$final)#0.7160323
corr_l_xy


```

2. Second, let’s fit the linear model that we will use to make predictions.

a. Use the function lm() to fit a linear model to summarize the relationship between X and Y
and store the output in an object called fit. Then, ask R to provide the contents of fit by
running its name. (R code only.) 

b. What is the fitted line? In other words, provide the formulâ Y =̂ α +̂ βX where you specify
each term (i.e., substitute Y for the name of the outcome variable, substitutê α for the
estimated value of the intercept coefficient, substitutê β for the estimated value of the slope
coefficient, and substitute X for the name of the predictor.

c. Create a visualization of the relationship between X and Y and add the fitted line to the
graph using the function abline()

```{r}
fit <- lm(final <- final ~ midterm, data = grades)
fit
summary(fit)


```
b)From the above result, the Fitted line is variable y(Final)-6 + .97variable x(Midterm)
Hence, Final =-6 + 0.97 Midterm


```{r}
grades <- read.csv("grades.csv")
library(tidyverse)
head(grades)

 student_scores <- ggplot(data = grades, mapping =aes(x=midterm,y=final))
 student_scores <- student_scores + geom_point()
 student_scores <- student_scores + geom_abline(method = "lm", se= FALSE, colours="blue", face ="bold")
 student_scores <- student_scores + xlab("midterm") + ylab("final")
 student_scores






```
3. Now, let’s use the fitted line to make some predictions.
a. Computinĝ Y based on X : Suppose that you earn 80 points in the midterm. What would
be your best guess of your predicted final exam score based on your performance in the
midterm? Please show your calculations and then answer the question with a full sentence
(including units of measurement). 

b. Computinĝ Y based on X : Now, suppose that you earn 90 points in the midterm. What
would be your best guess of your predicted final exam score based on your performance
in the midterm? Please show your calculations and then answer the question with a full
sentence (including units of measurement). 

c. Computing 4̂Y based on 4X : Suppose that you study a few extra hours and you earn 10
extra points in the midterm as a result. What would be your best guess of by how much
your predicted final exam score will change as a result of these 10 extra points in the
midterm? Please show your calculations and then answer the question with a full sentence
(including units of measurement)


###
a)Show workings
```{r}
-6.0059 + 0.9704 * 80 #computation for a





-6.0059 + 0.9704 * 90 #computation for b




0.9704 *10 #computation for c



```
From the above:a)My best guess using the assumed score of 80 in the midterm would be 71.63 approximately. The unit of measurement is 71.6 points.

b)My best guess using the assumed score of 90 in the midterm would be 81.3 The unit of measurement is 81.

c)My best guess from the 10 extra points would be a scale up by 9.704 on the final exam score.




4. What is the R2 of the fitted model? And, how would you interpret it? (Hint: the function cor()
might be helpful here.

```{r}
R2 <- cor(x = grades$midterm, y = grades$final)^2
R2
```
The R2 gives prediction according to the accuracy of the x and y variables. In this case it tells the performance of the fitted model, if the data perfectly fits or otherwise. From the results, the R2 of 0.51 which is 51% is acceptable, hence it explains that the final exam scores was predicted based on the midterm scores.